Ionization of biological molecules by multicharged ions by using the stoichiometric model

### 2019-09-23 NewMonti25.tex
new scale with ne fig

%\documentclass[pra,showpacs,twocolumn]{revtex4}
\documentclass[preprint]{revtex4}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{authblk}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\author{author}
%\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
%(CONICET-UBA).
%Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
%\author{author}
%\affil{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio (CONICET-UBA).
%Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
%\affil{Fac. de Ciencias Exactas y Naturales, Universidad de Buenos Aires}

@@

\begin{abstract}
\end{abstract}

@@

\tableofcontents
%\pacs{79.20.Rf, 68.49.Bc, 34.20.Cf,34.50.-s,34.35.+a}

@@

In the present study, we consider six atoms, $\alpha=$ H, C, N, O, P, 
and S, and six projectiles, antiprotons $\bar{p}$, H$^{+}$, He$^{+2}$, 
Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
Most of the organic molecules are composed of these atoms. Some 
particular molecules also include halogen atoms such as fluor and 
bromine; ionization cross sections of these elements have been 
previously published~\cite{miraglia2008,miraglia2009}.

The total ionization cross sections of these atoms were calculated using 
the CDW~[xxx]. The initial bound and final continuum radial wave 
functions were obtained by using the {\sc radialf} code, developed by 
Salvat and co--workers~\cite{salvat1995}, and a Hartree-Fock potential 
obtained from the Depurated Inversion Method~\cite{mendez2016,mendez2018}. 
We used a few thousand pivot points to solve the Schr\"{o}dinger 
equation, depending on the number of oscillations of the continuum 
state. The radial integration was performed using the cubic spline 
technique. The number of angular momenta considered varied from 8, at 
very low ejected-electron energies, up to 30, for the highest energies 
considered. The same number of azimuth angles were required to obtain 
the four-fold differential cross section. The calculation performed does 
not display prior-post discrepancies at all. Each atomic total cross 
section was calculated using 35 to 100 momentum transfer values, 28 
fixed electron angles, and around 45 electron energies depending on the 
projectile impact energy. In our theoretical treatment, we expand our 
final continuum wave function as usual,
\begin{equation}
\psi_{\overrightarrow{k}}^{-}(\overrightarrow{r})=\sum_{l=0}^{l_{\max
}}\sum_{m=-l}^{l}R_{kl}^{-}(r)Y_{l}^{m}(\widehat{r})Y_{l}^{m^{\ast }}
(\widehat{k})\,.
\label{eq:contwave}
\end{equation}
We are confident with our calculations up to $l_{\max}\sim 30$. Further 
details of the calculation are given in Ref.~\cite{montanari2017}. 
Simultaneously, we will be reporting state to state ionization cross 
sections for the 36 ion--target systems considered in the present 
work~\cite{miraglia2019}. We expect that these results will be useful
to estimate molecule fragmentation.

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig1.eps}
\includegraphics[width=0.9\textwidth]{figuras/atomicscaling.eps}
\caption{Reduced CDW total ionization cross section of six atomic 
targets. The curves are labeled with the charge state corresponding to 
the six multicharged projectiles.}
\label{fig:atomscaling}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig5.eps}
\includegraphics[width=0.9\textwidth]{figuras/ener_mean.eps}
\caption{Mean emitted energy distribution for ionization by impact of
multicharged ions, given by Eq.~(\ref{40}). Dashed lines for $\bar{p}$,
solid lines for ion charges $+1$, $+2$, $+4$, $+6$ and $+8$,
as indicated.}
\label{fig:emittedener}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig6.eps}
\includegraphics[width=0.9\textwidth]{figuras/ang_mean.eps}
\caption{Mean emitted angle distribution for ionization by impact of
multicharged ions.}
\label{fig:emittedang}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig2.eps}
\includegraphics[width=0.9\textwidth]{figuras/adn1.eps}
\caption{Reduced CDW ionization cross section by impact of multicharged 
ions. Experiments: \mbox{\Large$\circ$}~\cite{iriki2011} for proton 
impact and $\square$~\cite{rahman2016} for electron impact with 
equivelocity conversion.}
\label{fig:crossDNA_1}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig3.eps}
\includegraphics[width=0.9\textwidth]{figuras/adn2.eps}
\caption{Reduced CDW ionization cross section by impact of multicharged 
ions. Experiments: proton impact on $\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$
THF~\cite{wang2016}. Impact of $\ominus$ C$^{+4}$, 
$\boxminus$ C$^{+6}$, O$^{+6}$, F$^{+6}$, and
$\triangle\mkern-14mu-$ O$^{+8}$, F$^{+8}$ on 
uracil~\cite{agnihotri2012,agnihotri2013}. 
Symbols~$\rhd$~\cite{bug2017}, $\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} for electron impact with equivelocity 
conversion.}
\label{fig:crossDNA_2}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig4.eps}
\includegraphics[width=0.9\textwidth]{figuras/molscaling85.eps}
\caption{Scaled ionization cross section per weakly bound electron using
(a)~the Toburen numbers $\nu_{\alpha}^T$, and (b) our proposed numbers
$\nu_{\alpha}^{\text{CDW}}$. Experiments: proton impact on 
\mbox{\Large$\circ$} adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$ 
THF~\cite{wang2016}; electron impact on $\rhd$ pyrimidine~\cite{bug2017},
and $\lhd$, $\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:newscaling}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig_recta.eps}
\caption{Ionization cross sections by impact of protons at 0.5, 1 and 2 MeV
in terms of the number of active electrons given by Table~\ref{nn}.
Experiments: 
\mbox{\Large$\circ$}~adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014}, 
$\blacktriangle$ C2H$_7$N, CH$_5$N, methane and amonia~\cite{lynch1976},
\mbox{\scriptsize$\bigstar$} amonia and H$_2$~\cite{rudd1985}, and 
\mbox{\Large$\bullet$} water~\cite{luna2007}.}
\label{fig:recta}
\end{figure*}

@@

@@

The molecular binding energies of the valence electrons for adenine, 
cytosine, guanine, thymine and uracil are shown in Fig.~\ref{fig:bindener}. 
We can compute the center of gravity of the molecular energy levels as 
an average weighted by the electronic occupation number. The results 
obtained from the full molecular calculation are given in the first row
of Table~\ref{tab:gravener}. Similarly, we can compute the baricenter
of the energy levels defined by the Toburen rule and the one given by
Eq.~(\ref{eq:scalingCDW}), which are also given in 
Table~\ref{tab:gravener}. The average energy values obtained from the
Toburen rule are around 20\% of the full molecular ones. Surprisingly, 
the new scaling rule gives a significant 
improvement in the average energy values, reducing the relative errors
to about~7\%. This improvement would seem to indicate that the new 
scaling is appropiate.

@@

Furthermore, on the left side of the Fig.~\ref{fig:bindener}, we show 
the atomic Hartree--Fock energies of the constituent elements, which 
gives an insight on the distribution of the weakly bound electrons. 
A dashed line around $-11$~eV is drawn to separate the band in two. 
We can consider the atomic energy levels above this line as the ones
corresponding to the weakly bound electrons from Eq.~(\ref{eq:scalingCDW}).
For example, the $2s$ and $2p$ electrons of carbon are placed above
the separating line, which corresponds to the 4 electrons given by 
Eq.~(\ref{eq:scalingCDW}). In the case of O, only the 4 electrons of 
the $2p$ orbitals are located above the separating line.

The electrons in the molecular 
targets corresponding to the $2s$ of N and C in the atomic case are 
placed in a secondary gap. The lower levels of the secondary gap 
correspond to the oxygen. The number of electrons in the valence 
shell for all O would seem to be equal to 4, as previously setted by the 
scaling rule Eq.~(\ref{eq:scalingCDW}). The N case is not as 
straightfoward; the electrons that would correspond to the $2s$ are 
located just below the line. The $\nu_{\alpha }^{\text{CDW}}$ value 
given for N would suggest that a significant amount of electron density 
from the secondary gap is shared with the valence band.

@@

Consider that the total amount of electrons $Q_{\alpha }$ on the element
$\alpha$ are equally distributed on all the $\alpha$ atoms. Therefore, 
each element $\alpha$ will have an additional charge: 
$q_{\alpha}=Q_{\alpha}/n_{\alpha}$, which can be positive or negative.
This value will depend on the relative electronegativity value with 
respect to the other atoms~\cite{rappe1991}. Now, instead of an
integer number of elements $n_{\alpha}$ of the atom $\alpha$, we have a 
fractional number of atoms given by 
\begin{equation}
n_{\alpha }^{\prime }=n_{\alpha }-
\frac{q_{\alpha }}{\nu_{\alpha }^{\text{CDW}}}
\label{eq:newstoi}
\end{equation}%
In the case of neutral atoms, $q_{\alpha}=0$ and we recover 
$n_{\alpha}^{\prime}=n_{\alpha}$, as it should be.

In Table~\ref{tab:newstoi}, we display the charge $q_{\alpha}$ of four 
DNA molecules, obtained from the full molecular calculation.

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.12\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}|p{
0.07\textwidth}|p{0.07\textwidth}|p{0.25\textwidth}|}
\hline
Element & C & H & N & O & New stoichiometry \\
\hline
Adenine & +0.32 & +0.23 & --0.55 &       & 
C$_{4.92}$H$_{4.77}$N$_{5.14}$ \\ 
\hline
Cytosine & +0.28 & +0.21 & --0.56 & --0.53 & 
C$_{3.93}$H$_{2.79}$N$_{5.14}$O$_{1.13}$ \\ 
\hline
Guanine & +0.46 & +0.20 & --0.58 & --0.36 & 
C$_{4.89}$H$_{4.80}$N$_{5.15}$O$_{1.09}$ \\ 
\hline
Thymine & +0.20 & +0.19 & --0.54 & --0.52 & 
C$_{4.95}$H$_{1.95}$N$_{6.13}$O$_{2.13}$ \\ 
\hline
\end{tabular}
\caption{Effective charge $q_{\alpha}$ and new stoichiometric formula 
defined by Eq.~(\ref{eq:newstoi}) for four DNA molecules.}
\label{tab:newstoi}
\end{center}
\end{table}

By implementing Eq.~(\ref{eq:newstoi}), it is possible to determine a 
new stoichiometric formula (last column of Table~\ref{tab:newstoi}). 
Now, instead of having an integer number of atoms $n_{\alpha}$, we obtain 
a fractional number $n_{\alpha}^{\prime}$. 
Molecular cross sections $\sigma^{\prime}$ can be computed considering 
such values. Relative errors for the ionization cross sections were 
computed for the DNA bases from Table~\ref{tab:newstoi}. The differences 
obtained were of just very few percents, which indicates that the modified 
SSM is a quite robust model to handle these type of molecule within 
the range error expected for this model.

@@

We calculated ionization cross sections by impact of antiprotons, 
H$^{+}$, He$^{+2}$, Be$^{+4}$, C$^{+6}$, and O$^{+8}$ for seventeen 
molecules containing H, C, N, O, P and S with the CDW method. The 
importance of the influence of $Z$ was observed in the mean energy 
$\overline{E}_{\alpha}$ and angle $\overline{\theta}_{\alpha}$. 
For a given target $\alpha$, as the nuclear charge of the projectile 
$Z$ increases, so does the mean energy of emittion $\overline{E}_{\alpha}$. 
Conversely, the mean angle of emission  $\overline{\theta}_{\alpha}$ 
decreases. At high impact energy, say larger than 1 MeV/amu, these 
values converge to the Born approximation, which embodies the simple 
$Z^{2}$ law. The seventeen molecules selected were investigated using 
the simple stoichiometric model. Results for eigth DNA bases were 
presented and compared with the sparse available experiments. We explore 
the rule of Toburen which scales all the molecular ionization cross 
section when divided by the number of weakly bound valence electrons 
$\nu_{\alpha}^T$ given by Eq.~(\ref{27}). 
We found the rule scales much better when normalizing our theoretical 
ionization cross sections to the number $\nu_{\alpha}^{\text{CDW}}$ 
given by Eq.~(\ref{eq:scalingCDW}). Finally, we attempt to improve the 
stoichiometric model by using the Mulliken charge to define a new model 
containing fractional rather than integer proportions. No substantial 
correction was found, which indicates that our modified SSM works quite 
well. By inspecting the molecular binding energy from quantum mechanical 
structure calculations, we were able to understand the values defined
for $\nu_{\alpha}^{\text{CDW}}$.

The main objective of this article is to provide the tools to estimate 
any inelastic parameter parameter --such as emission angle, emitted mean 
energy and cross section-- by the impact of any multicharged on any 
molecule containing H, C, N, O, P and S, with the help of the 
stoichimetrical model. Our goal was quite ambitious, considering the 
simplicity of our proposal. However, we think our results could be used 
to estimate the ionization magnitude with an acceptable level of 
uncertainty.

@@

@@

@@

\bibitem{pimblott2007} 
S.M. Pimblott and J. A. LaVerne, Radiation
Physics and Chemistry 76, 1244-1247 (2007)

\bibitem{clementi}
E. Clementi, C. Roetti,
At. Data Nucl. Data Tables 14, 177--478 (1974).

>>>>

%\documentclass[pra,showpacs,twocolumn]{revtex4}
\documentclass[preprint]{revtex4-2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{gensymb}
\usepackage{float}
\usepackage{MnSymbol}
\usepackage{gensymb}

@@

\author{author}
\address{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET-UBA).
Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
\author{author}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio (CONICET-UBA).
Casilla de correo 67, sucursal 28 (C1428EGA) Buenos Aires, Argentina.}
\affiliation{Fac. de Ciencias Exactas y Naturales, Universidad de Buenos Aires}

@@

\begin{abstract}
In this work we investigate the ionization of molecules of biological 
interest by impact of multicharged ions in the intermediate to high 
energy range. We performed full non--perturbative distorted--wave 
calculations (CDW) for thirty six collisional systems, six charged projeciles 
(antiprotons, H, He, B, C and O) and by six atomic targets H, C, N, O, 
F and S. These atoms are the constituents of most of the DNA and 
biological molecules. By considering the stoichiometric composition of 
the different molecules we approximate the ionization cross sections 
as linear combination of the atomic ones.  We show that this simple 
stoichiometric model (SSM) gives reasonable good results for complex 
molecules. We focused on seventeen molecules including the DNA 
(adenine, cytosine, guanine, thymine, uracil), DNA backbone, the 
pyrimidines, tetrahydrofuran (THF), and the C$_n$H$_n$ compounds. The 
full comparison of 
the total ionization cross sections for the different ions normalized 
to the square of the ion charge gives an interesting picture of the
experimental state of art. The energy and angular distributions of the 
emitted electrons are also inspected due to its relation with the 
radiation damage. Our results show that these two parameters are highly 
dependent on the projectile charge state, which cannot be noted in a
first order perturbative calculation. We also analyze the extensively 
used Toburen 
scaling of the total ionization cross sections of molecules with the 
number of weakly bound electrons. We tested this scaling for the seventeen 
different molecules and the six projectiles per each, which allows to 
have an insight on the limitation of this scaling. Instead, based on 
the present CDW results for the atoms, we propose a different number 
of electrons per atom. This CDW--based scaling highly improves the 
single curve behaviour for all the targets and ions studied here, in the 
intermediate to high energy region. The new scaling also describes well 
the available experimental data, including small molecules such as H2, 
water, methane and ammonia. For completeness, we performed full 
molecular calculations for the five DNA molecules, and tested a 
corrected stoichiometric formula based on the actual number of 
electrons around each molecular center. The difference of these more 
realistic ionization cross sections and the SSM ones was less than 
$3\%$, being an indication of the reliability and robustness of the 
SSM to deal with this type of molecules. Based on the extensive 
ion-target systems included in the present study, and the results 
obtained, the possibilities of the SSM and the new scaling are of 
great impact.
\end{abstract}

@@

%\tableofcontents
%\pacs{79.20.Rf, 68.49.Bc, 34.20.Cf,34.50.-s,34.35.+a}

@@

In the present study, we consider six atoms, $\alpha=$ H, C, N, O, P, 
and S, and six projectiles, antiprotons $\bar{p}$, H$^{+}$, He$^{+2}$, 
Be$^{+4}$, C$^{+6}$, and O$^{+8}$. 
Most of the organic molecules are composed of these atoms. Some 
particular molecules also include halogen atoms such as fluor and 
bromine; ionization cross sections of these elements have been 
previously published~\cite{miraglia2008}.

The total ionization cross sections of these atoms were calculated using 
the CDW~\cite{miraglia2008,fainstein1988,miraglia2009}. The initial bound and final continuum radial wave 
functions were obtained by using the {\sc radialf} code, developed by 
Salvat and co--workers~\cite{salvat1995}, and a Hartree-Fock potential 
obtained from the Depurated Inversion Method~\cite{mendez2016,mendez2018}. 
We used a few thousand pivot points to solve the Schr\"{o}dinger 
equation, depending on the number of oscillations of the continuum 
state. The radial integration was performed using the cubic spline 
technique. The number of angular momenta considered varied from 8, at 
very low ejected-electron energies, up to 30, for the highest energies 
considered. The same number of azimuth angles were required to obtain 
the four-fold differential cross section. The calculation performed does 
not display prior-post discrepancies at all. Each atomic total cross 
section was calculated using 35 to 100 momentum transfer values, 28 
fixed electron angles, and around 45 electron energies depending on the 
projectile impact energy. In our theoretical treatment, we expand our 
final continuum wave function as usual,
\begin{equation}
\psi_{\overrightarrow{k}}^{-}(\overrightarrow{r})=\sum_{l=0}^{l_{\max
}}\sum_{m=-l}^{l}R_{kl}^{-}(r)Y_{l}^{m}(\widehat{r})Y_{l}^{m^{\ast }}
(\widehat{k})\,.
\label{eq:contwave}
\end{equation}
We are confident with our calculations up to $l_{\max}\sim 30$. Further 
details of the calculation are given in Ref.~\cite{montanari2017}. 
Simultaneously, we will be reporting state to state ionization cross 
sections for the 36 ion--target systems considered in the present 
work~\cite{miraglia2019}. A great numerical effort was paid to obtain 
the present results, and we expect that they will be useful to estimate 
molecule fragmentation.

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/Fig1.eps}
%\includegraphics[width=0.9\textwidth]{figuras/atomicscaling.eps}
\caption{Reduced CDW total ionization cross section of six atomic 
targets. The curves are labeled with the charge state corresponding to 
the six multicharged projectiles.}
\label{fig:atomscaling}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig5.eps}
%\includegraphics[width=0.9\textwidth]{figuras/ener_mean.eps}
\caption{Mean emitted energy distribution for ionization by impact of
multicharged ions, given by Eq.~(\ref{40}). Dashed lines for $\bar{p}$,
solid lines for ion charges $+1$, $+2$, $+4$, $+6$ and $+8$,
as indicated.}
\label{fig:emittedener}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig6.eps}
%\includegraphics[width=0.9\textwidth]{figuras/ang_mean.eps}
\caption{Mean emitted angle distribution for ionization by impact of
multicharged ions.}
\label{fig:emittedang}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig2.eps}
%\includegraphics[width=0.9\textwidth]{figuras/adn1.eps}
\caption{Reduced CDW ionization cross section by impact of multicharged 
ions. Experiments: \mbox{\Large$\circ$}~\cite{iriki2011} for proton 
impact and $\square$~\cite{rahman2016} for electron impact with 
equivelocity conversion.}
\label{fig:crossDNA_1}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig3.eps}
%\includegraphics[width=0.9\textwidth]{figuras/adn2.eps}
\caption{Reduced CDW ionization cross section by impact of multicharged 
ions. Experiments: proton impact on $\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$
THF~\cite{wang2016}. Impact of $\ominus$ C$^{+4}$, 
$\boxminus$ C$^{+6}$, O$^{+6}$, F$^{+6}$, and
$\triangle\mkern-14mu-$ O$^{+8}$, F$^{+8}$ on 
uracil~\cite{agnihotri2012,agnihotri2013}. 
Symbols~$\rhd$~\cite{bug2017}, $\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} for electron impact with equivelocity 
conversion.}
\label{fig:crossDNA_2}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig4.eps}
%\includegraphics[width=0.9\textwidth]{figuras/molscaling85.eps}
\caption{Scaled ionization cross section per weakly bound electron using
(a)~the Toburen numbers $\nu_{\alpha}^T$, and (b) our proposed numbers
$\nu_{\alpha}^{\text{CDW}}$. Experiments: proton impact on 
\mbox{\Large$\circ$} adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014} and $\meddiamond$ 
THF~\cite{wang2016}; electron impact on $\rhd$ pyrimidine~\cite{bug2017},
and $\lhd$, $\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:newscaling}
\end{figure*}

@@

\begin{figure*}[t!]
\centering
%\includegraphics[width=0.75\textwidth]{figuras/Fig_finales/fig_recta.eps}
\includegraphics[width=0.75\textwidth]{figuras/scale_ne.eps}
\caption{Ionization cross sections by impact of protons at 0.5, 1 and 2 MeV
in terms of the number of active electrons given by Table~\ref{nn}.
Experiments: 
\mbox{\Large$\circ$}~adenine~\cite{iriki2011}, 
$\triangle$ uracil~\cite{itoh2013}, 
$\bigtriangledown$ pyrimidine~\cite{wolff2014}, 
$\blacktriangle$ C2H$_7$N, CH$_5$N, methane and amonia~\cite{lynch1976},
\mbox{\scriptsize$\bigstar$} amonia and H$_2$~\cite{rudd1985}, and 
\mbox{\Large$\bullet$} water~\cite{luna2007}.}
\label{fig:recta}
\end{figure*}

@@

Finishing the present work we noted an accepted manuscript by 
Luedde~{\it et al.} \cite{luedde2019} who also noted a similar scaling 
at least for C, N, and O and reinforces the present finding.

@@

The molecular binding energies of the valence electrons for adenine, 
cytosine, guanine, thymine and uracil are shown in Fig.~\ref{fig:bindener}. 
We can compute the center of gravity of these molecular energy levels as 
an average weighted by the electronic occupation number. The results 
obtained from the full molecular calculation are given in the first row
of Table~\ref{tab:gravener}. Similarly, we can compute the baricenter
of the energy levels defined by the Toburen rule and the new rule, given by
Eq.~(\ref{eq:scalingCDW}), which are also given in 
Table~\ref{tab:gravener}. The average energy values obtained from the
Toburen rule are around 20\% of the full molecular ones. Surprisingly, 
the new scaling rule gives a significant 
improvement in the average energy values, reducing the relative errors
to about~7\%. This improvement would seem to indicate that the new 
scaling is appropiate.

@@

Furthermore, on the left side of the Fig.~\ref{fig:bindener}, we show 
the atomic Hartree--Fock energies of the constituent elements, which 
gives an insight on the distribution of the weakly bound electrons. 
A dashed line around $-11$~eV is drawn to separate the band in two. 
We can consider the atomic energy levels above this line as the ones
corresponding to the weakly bound electrons from Eq.~(\ref{eq:scalingCDW}).
For example, the $2s$ and $2p$ electrons of carbon are placed above
the separating line, which corresponds to the 4 electrons given by 
Eq.~(\ref{eq:scalingCDW}). In the case of O, only the 4 electrons of 
the $2p$ orbitals are located above the separating line, which 
corresponds to the number of weakly bound electron given by 
our new scaling. 
The N case is not as straightfoward; the $\nu_{\alpha }^{\text{CDW}}$ 
value given for N would suggest that some $2s$ electrons corresponding 
to the molecular squeme are also weakly bonded, leading to 4 electrons.

\subsubsection{A modified stoichiometric model}

@@

Consider that the total amount of electrons $Q_{\alpha }$ on the element
$\alpha$ are equally distributed on all the $\alpha$ atoms. Therefore, 
each element $\alpha$ will have an additional charge: 
$q_{\alpha}=Q_{\alpha}/n_{\alpha}$, which can be positive or negative.
This value will depend on the relative electronegativity value with 
respect to the other atoms~\cite{rappe1991}. Now, instead of an
integer number of elements $n_{\alpha}$ of the atom $\alpha$, we have a 
fractional number of atoms given by 
\begin{equation}
n_{\alpha }^{\prime }=n_{\alpha }-
\frac{q_{\alpha }}{\nu_{\alpha }^{\text{CDW}}}
\label{eq:newstoi}
\end{equation}%
In the case of neutral atoms, $q_{\alpha}=0$ and we recover 
$n_{\alpha}^{\prime}=n_{\alpha}$, as it should be. 
In Table~\ref{tab:newstoi}, we display the average effective charge 
per atom $q_{\alpha}$ of C, H, N and O, for five DNA molecules. 
The charges were obtained from the full molecular calculation described 
above.

\begin{table}[H]
\begin{center}
\begin{tabular}{|p{0.12\textwidth}|p{0.07\textwidth}|p{0.07\textwidth}|p{
0.07\textwidth}|p{0.07\textwidth}|p{0.25\textwidth}|}
\hline
Element & C & H & N & O & New stoichiometry \\
\hline
Adenine & +0.32 & +0.23 & --0.55 &       & 
C$_{4.92}$H$_{4.77}$N$_{5.14}$ \\ 
\hline
Cytosine & +0.28 & +0.21 & --0.56 & --0.53 & 
C$_{3.93}$H$_{2.79}$N$_{5.14}$O$_{1.13}$ \\ 
\hline
Guanine & +0.46 & +0.20 & --0.58 & --0.36 & 
C$_{4.89}$H$_{4.80}$N$_{5.15}$O$_{1.09}$ \\ 
\hline
Thymine & +0.20 & +0.19 & --0.54 & --0.52 & 
C$_{4.95}$H$_{1.81}$N$_{6.13}$O$_{2.13}$ \\ 
\hline
Uracil & +0.31 & +0.22 & --0.59 & --0.47 & 
C$_{3.92}$H$_{1.78}$N$_{4.15}$O$_{2.12}$ \\ 
\hline
\end{tabular}
\caption{Average effective Mulliken charge per atom $q_{\alpha}$, and 
new stoichiometric formula defined by Eq.~(\ref{eq:newstoi}) for five 
DNA molecules.}
\label{tab:newstoi}
\end{center}
\end{table}

By implementing Eq.~(\ref{eq:newstoi}), it is possible to determine a 
new stoichiometric formula (last column of Table~\ref{tab:newstoi}). 
Now, instead of having an integer number of atoms $n_{\alpha}$, we obtain 
a fractional number $n_{\alpha}^{\prime}$. New molecular cross sections 
$\sigma^{\prime}_{M}=\sum_{\alpha}n_{\alpha}'\sigma_{\alpha}$ can be 
computed considering 
such values. Relative errors for the ionization cross sections were 
computed for the DNA bases from Table~\ref{tab:newstoi}. The differences 
obtained were less than 3\%, which indicates that the modified 
SSM is a quite robust model to handle these type of molecules within 
the range error expected for this model.

@@

In this work we have encouraged the calculation of ionization cross 
sections of seventeen biological molecules containing H, C, N, O, P and 
S by impact of antiprotons, H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, 
and O$^{8+}$. 
To that end we have employed the full CDW method including... and a 
simple stoichiometric model. 
The mean energy and angle of the emitted electrons, of importance in 
post-collisional radiation damage,  has also been calculated, showing 
a clear dependence with the ion charge $Z$. For a given target as $Z$ 
increases, $\overline{E}_{\alpha}$
increases, but $\overline{\theta}_{\alpha}$ decreases. At impact
energies larger than 1 MeV/amu, these values converge to the Born
approximation, which embodies the simple $Z^{2}$ law.

Total ionization cross sections for adenine, cytocine, thymine, gaunine, 
uracil, DNA backbone, pyrimidine and THF are presented and compared 
with the the scarse available experiments. We explore the rule of 
Toburen which scales
all the molecular ionization cross section normalizing with certain 
number of
weakly bound or valence electrons. We found that ionization cross 
sections scales much better when normalizing with the number of active 
electrons in the collision obtained from the CDW results for atoms. 
This new scaling was tested with very good results for the six 
projectiles and seventeen molecules studied here. The comparison with 
the experimental data reinforce this results. Moreover, we also test 
the scaling by including experimental data of ionization of H$_2$, 
water, methane and ammonia by proton impact showing good agreement at
intermediate to high energies.
Finally, we performed full molecular calculations for the DNA basis. 
By inspecting the molecular binding energy from quantum mechanical
structure calculations, we were able to understand the number of 
electrons proposed in our new CDW-based scaling. We attempt to improve 
the stoichiometric model by using the Mulliken charge to get fractional
rather than integer proportions. It is interesting to remark that no 
substantial
correction was found, which indicates that the SSM works quite well.

The main objective of this article was to provide the tools to estimate 
any inelastic parameter parameter --such as emission angle, emitted mean
energy and cross section-- by the impact of any multicharged projectiles 
on any molecule containing H, C, N, O, P and S, with the help of the
stoichimetrical model. Our goal was quite ambitious, considering the
simplicity of our proposal. However, the present results show the 
validity of the simple stoichiometric model to estimate the ionization 
magnitude with an acceptable level of uncertainty.

@@

\bibitem{fainstein1988}
Fainstein P.D., Ponce V. H. and Rivarola R. D. 
J. Phys. B: At. Mol. Opt. Phys. \textbf{21} 287 (1988).

@@

\bibitem{luedde2019}
H J Luedde et al 2019 
J. Phys. B: At. Mol. Opt. Phys. in press https://doi.org/10.1088/1361-6455/ab3a63

@@

%\bibitem{pimblott2007} 
%S.M. Pimblott and J. A. LaVerne, Radiation
%Physics and Chemistry 76, 1244-1247 (2007)

%\bibitem{clementi}
%E. Clementi, C. Roetti,
%At. Data Nucl. Data Tables 14, 177--478 (1974).
